---
title: "Group 4 Report"
author: "Richard Davies, Katarina Lundervol, Pakwanja Desiree Twea"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
Requred package: tinytex, tidyverse, here, rstatix, corrplot, ggplot2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TidyR data setup

```{r include=FALSE}
library(tinytex)
library(tidyverse)
library(here)
library(rstatix)
library(corrplot)
library(ggplot2)
```


The data is composed of two datasets with other lapping patients (see ID), exam.dataset.txt consists of PCR rdts for COVID19 during the 2020 pandemic, data includes information on patients including clinic, gender, test results week of pandemic, and exam_joindata.txt containing endpoint titer data from some of the patients (information can be found in the files codebook_exam_data.html, exam.descr.md). Data files were joined and processed as per conventions of tidyR for later visulization and analysis.

```{r include=FALSE}
test_join <- read_delim("data/exam_joindata.txt", col_names = TRUE) %>%
  rename(ID = id)

complete_data <- read_delim("data/exam_dataset.txt", col_names = TRUE) %>%
  separate(col = "subject", into = c("ID", "first_name", "last_name"), sep = " ") %>%
  separate(col = "gender-age", into = c("gender", "age"), sep = "-") %>%
  distinct(.keep_all = FALSE) %>%
  rename("test_id" = "1_test_id", "pan_day" = "pan day", "time_measurement" = "time measurement", "value" = ".value") %>%
  pivot_wider(names_from = time_measurement, values_from = value) %>%
  mutate(ID = as.numeric(ID),
         gender = as.factor(gender),
         age = as.numeric(age),
         clinic_name = as.factor(clinic_name),
         result = as.factor(result),
         demo_group = as.factor(demo_group),
         drive_thru_ind = as.factor(drive_thru_ind),
         ct_result = as.numeric(ct_result),
         orderset = as.numeric(orderset),
         payor_group = as.factor(payor_group),
         patient_class = as.factor(patient_class),
         pan_day = as.numeric(pan_day),
         test_id = as.factor(test_id),
         row = as.numeric(row),
         rec_ver_tat = as.numeric(rec_ver_tat),
         col_rec_tat = as.numeric(col_rec_tat)
  ) %>%
  select(-row, -test_id, -demo_group) %>%
  arrange(ID) %>%
  mutate(rvt = if_else(rec_ver_tat > 100, "High", "Low")) %>%
  mutate(pan_week = pan_day / 7) %>%
  mutate(dti_yes_no = if_else(drive_thru_ind == 1, "Yes", "No")) %>%
  mutate(ct_orderset = ct_result * orderset) %>%
  select(ID, age, gender, everything()) %>%
  full_join(test_join, join_by(ID))

```



An overview of the data follows 

```{r}
summary(complete_data)
```


```{r}
naniar::gg_miss_var(complete_data)
```




## Descriptive plots

Visulization was done using a number of plots. Significant correlations are observiable with age, and orderset and days since the pandemic. Indicating early in the pandemic individuals seen in the clinic were usually young adults and children  

```{r, echo=FALSE}
numeric_data <- complete_data %>% select(where(is.numeric))
numeric_data <- numeric_data[, c(-1, -8, -9)] # ID, ct_orderset, pan_week removed due to inderpendence) 

correlation <- round(cor(numeric_data, use = "complete.obs"), 1)
```



```{r, echo=FALSE}
cor_pmat(numeric_data)
```

```{r, echo=FALSE}
correlation <- round(cor(numeric_data, use = "complete.obs"), 1) # Makes object 
corrplot(correlation)
```
```{r, echo = FALSE}
ggplot(data = numeric_data, aes(y = age, x = pan_day)) +
  geom_point()
```
Suprisingly over the course of the pandemic there was no indication that the wait time from collection of patient sample to PCR test result was reduced.

```{r include=FALSE}
wait_stats <- complete_data %>%
  select(pan_week, rec_ver_tat) %>%
  mutate(pan_week = round(pan_week)) %>%
  group_by(pan_week) %>%
  summarize(mean_value = mean(rec_ver_tat), std = sd(rec_ver_tat), median(rec_ver_tat))

```


```{r, echo =FALSE}
wait_stats
```

```{r, echo=FALSE}
ggplot(data = complete_data, aes(x = pan_week, y = rec_ver_tat)) +
  geom_point() +
  geom_smooth(method = "lm")
```

We observe that there are slightly more females than males in the cleaned data set. The bar plot below indicates that more people from both genders took the test at the drive-through than those who did not. However, from observation, the difference between genders is slightly larger among those who took the test at the drive-through compared to those who did not. To examine the relationship between gender and drive-through test participation, we conducted a Chi-squared test. With a p-value of 0.8728, we fail to reject the null hypothesis, suggesting that there is no significant association between gender and drive-through test participation in this data set.
```{r}
dt_and_gender <- table(complete_data$gender, complete_data$drive_thru_ind)
# Create dt_and_gender as a data frame from the table
dt_and_gender_df <- as.data.frame(dt_and_gender)
print(chisq.test(dt_and_gender))

# Rename columns for clarity (since table() creates V1, V2, and Freq by default)
colnames(dt_and_gender_df) <- c("Gender", "Drive_Through", "Count")

# Now use ggplot2 to plot
ggplot(dt_and_gender_df, aes(x = Drive_Through, y = Count, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Gender Distribution of Drive-Through Test Participants",
       x = "Drive-Through Indicator", 
       y = "Count") +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4")) +  # Custom colors
  theme_minimal()
```

A comparison of cycle threshold values between men and women shows a similar distribution in the number of reaction cycles. However, males showed the lowest cycle threshold value compared to females. Despite this, the majority of threshold values exceeded 40 cycles for both genders. The p-value of 0.272 from the one-way ANOVA test means that we fail to reject the null hypothesis, suggesting that gender does not significantly affect the cycle threshold value.

```{r}
ct_result_cat <- cut(complete_data$ct_result, c(10, 20, 30, 40, 50))
summary(ct_result_cat)

ct_result_dist <- complete_data %>%
  group_by(gender) %>%
  count(ct_result)

ggplot(ct_result_dist, 
       aes(
         x = 1:nrow(ct_result_dist),
         y = ct_result,
         group = gender,
         color = gender)
) + 
  geom_point() +
  facet_wrap(vars(gender))

ct_result_gender_anova <- aov(complete_data$ct_result ~ complete_data$gender, data = complete_data)
summary(ct_result_gender_anova)
```

The distribution of patients by payor group shows that the most common groups were commercial and government, with a significant proportion of patients listed as NA or unassigned. Despite this, a similar pattern was observed between government and commercial patients, while for charity care, medical assistance, other, and self-pay patients the minimum cycle threshold value was higher compared to the former groups. The one-way ANOVA test result is highly significant, indicating that the cycle threshold value is influenced by the payor group.

```{r}
ct_result_dist_pg <- complete_data %>%
  group_by(payor_group) %>%
  count(ct_result)

ggplot(
  ct_result_dist_pg,
  aes(
    x = 1:nrow(ct_result_dist_pg),
    y = ct_result,
    group = payor_group,
    color = payor_group
    )
  ) +
  geom_point() +
  facet_wrap(vars(payor_group))

ct_result_pg_anova <- aov(complete_data$ct_result ~ complete_data$payor_group, data = complete_data)
summary(ct_result_pg_anova)
```

The distribution of results by gender shows a similar pattern for both, with women comprising a slightly larger proportion in each result category except for the invalid category. The chi-square test yielded a p-value of 0.5481, meaning that we fail to reject the null hypothesis, suggesting that gender does not have a significant influence on the test result.

```{r}
tests_by_gender <- complete_data %>%
  group_by(gender) %>%
  count(result)

tests_by_gender

ggplot(data = tests_by_gender, aes(x = result, y = n, fill = gender)) +
  geom_bar(stat = "identity", color="black", position=position_dodge())+
  theme_minimal() +
  scale_fill_manual(values=c('#999999','#E69F00')) +
# Use brewer color palettes
  scale_fill_brewer(palette="Blues")

test_and_gender <- table(complete_data$gender, complete_data$result)
test_and_gender

#chi-square test
test_and_gender_df <- as.data.frame(test_and_gender)
print(chisq.test(test_and_gender))

```


## Statistics

wilcox test indicated that the more tests from a single patient does not associate with increase positive test, in fact the opposite trend was noted.

```{r include=FALSE}
positive_visits <- complete_data %>%
  mutate(positive = if_else(result == "positive", 1, 0)) %>%
  group_by(ID) %>%
  count(positive, ID)

positive_visits <- positive_visits[, -1]
positive_visits$positive <- factor(positive_visits$positive)
levels(positive_visits$positive) <- c("negative", "positive")

positive_visits %>% 
  group_by(positive) %>%
  summarize(mean(n), median(n))
```


```{r, , echo=FALSE}
positive_visits %>% 
  wilcox.test(n ~ positive, data = .) %>%
  broom::tidy()
```

While drive-through use was associated with an increase in positive tests


```{r, , echo=FALSE}
chisq.test(x = positive_visits$n, y = positive_visits$positive, simulate.p.value = TRUE)
```


